BoRA: Bayesian Hierarchical Low-Rank Adaption for Multi-task Large Language Models

### 2024-10-19 paper/main.tex
submission state (pre review changes)

%\documentclass{article}
\documentclass[fullpaper]{nldl}
\paperID{42}
\vol{V}
\usepackage{graphicx} % Required for inserting images
\usepackage{authblk}

\title{BoRA: Bayesian Hierarchical  Low-Rank Adaption for Multi-task Large Language Models}
\author[1]{Simen Eide}
\author[2]{Arnoldo Frigessi}
\affil[1]{University of Oslo, Schibsted}
\affil[1]{University of Oslo}
\date{March 2024}

%\usepackage{biblatex} % [backend=biber,style=alphabetic,sorting=ynt] %not necessary when using template
\addbibresource{references.bib}
\usepackage{amsmath}
\usepackage{amssymb}
\newcommand{\R}{\mathbb{R}}
\usepackage{booktabs}
%PACKAGES FROM TEMPLATE THAT SEEMS NECESSARY
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{listings}
\lstset{
  basicstyle=\small\ttfamily,
  breaklines,
}
\usepackage{enumitem}
\usepackage{algorithm}
\usepackage{algorithmic}
\newcommand{\theHalgorithm}{\arabic{algorithm}}
%%% END OF PACKAGES FROM TEMPLATE

@@

In these multi-task cases, the practitioner can either (1) train one model for each task and dataset separately, or (2) train one model for all tasks and condition the specific output wanted (e.g \cite{Raffel2019}).
There is a trade-off between these two approaches: Train each task separately and the model will be able to specialize on each task, but it may suffer when limited data are available for some task. Train one model for all tasks, and the model will be able to share information between tasks, but it may not be able to specialize on each task due to limited model capacity.

In this paper we suggest a way to circumvent these trade-offs and presents a generalization of the two LoRA fine-tuning methods for multi-task problems, which we call Bayesian Hierarchical Low-Rank Adaption method (BoRA).
The BoRA method allows tasks to share information between each other through sharing a set of global hierarchical prior parameters, in a Bayesian hierarchical setting. 
This means that tasks with little data can benefit from the global structure across related tasks, while tasks with more data can specialize on their specific task.
We show in Section \ref{sec:BoRA} that BoRA can be seen as a generalization of the two approaches above.

>>>>

%\documentclass{article}
\documentclass[fullpaper]{nldl}
\paperID{1}
\vol{V}
\usepackage{graphicx} % Required for inserting images
\usepackage{authblk}

\title{BoRA: Bayesian Hierarchical  Low-Rank Adaption for Multi-task Large Language Models}
\author[1]{Simen Eide}
\author[2]{Arnoldo Frigessi}
\affil[1]{University of Oslo, Schibsted}
\affil[1]{University of Oslo}
\date{July 2024}

\usepackage{biblatex}[backend=biber,style=alphabetic,sorting=ynt] %not necessary when using template
\addbibresource{references.bib}
\usepackage{amsmath}
\usepackage{amssymb}
\newcommand{\R}{\mathbb{R}}
\usepackage{booktabs}
%PACKAGES FROM TEMPLATE THAT SEEMS NECESSARY
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{listings}
\lstset{
  basicstyle=\small\ttfamily,
  breaklines,
}
\usepackage{enumitem}
\usepackage{algorithm}
\usepackage{algorithmic}
\newcommand{\theHalgorithm}{\arabic{algorithm}}
%%% END OF PACKAGES FROM TEMPLATE

@@

In these multi-task cases, the practitioner can either (1) train one model for each task and dataset separately, or (2) train one model for all tasks and condition the specific output wanted (e.g~\cite{Raffel2019}).
There is a trade-off between these two approaches: Train each task separately and the model will be able to specialize on each task, but it may suffer when limited data are available for some task. Train one model for all tasks, and the model will be able to share information between tasks, but it may not be able to specialize on each task due to limited model capacity.

In this paper we suggest a way to circumvent these trade-offs and presents a generalization of the two LoRA fine-tuning methods for multi-task problems, which we call Bayesian Hierarchical Low-Rank Adaption method (BoRA).
The BoRA method allows tasks to share information between each other through sharing a set of global hierarchical prior parameters, in a Bayesian hierarchical setting. 
This means that tasks with little data can benefit from the global structure across related tasks, while tasks with more data can specialize on their specific task.
We show in Section~\ref{sec:BoRA} that BoRA can be seen as a generalization of the two approaches above.
